Global study of well-being and internet adoption | Online analysis supplement

### 2021-11-28 03-model-estimate.Rmd
Include gwp in models

```{r}
itu <- read_rds("data/itu.rds")
gbd <- read_rds("data/gbd.rds") # 10 to 24 year olds
```

@@

```{r data-prepare}
# Merge predictors to outcomes
dat <- left_join(gbd, itu)

# Center and scale year, and contrast code factors
dat <- dat %>% 
  mutate(year = (year - 2010) / 10)
dat$sex <- factor(dat$sex)
contrasts(dat$sex) <- "contr.sum"
dat$age <- factor(dat$age)
contrasts(dat$age) <- "contr.sum"
```

@@

@@

```{r}
dat_model1 <- dat %>% 
  # Transform edge proportions to fit beta distribution
  mutate(internet = if_else(internet == 0, .0000001, internet)) %>% 
  mutate(internet = if_else(internet == 1, .9999999, internet)) %>% 
  # Indicate unique ITU values for MV model...
  arrange(region, country, year, cause, sex, age) %>% 
  # ...Use only the first internet value within a year-country
  group_by(region, country, year) %>% 
  mutate(itu = 1:n() == 1) %>% 
  ungroup()
```

@@

```{r}
# Define models of outcomes
bf_1 <- bf(
  val ~ 
    year * sex + year * age + 
    (year * sex + year * age |c| country) +
    (year * sex + year * age |r| region),
  family = gaussian()
)

bf_1_se <- bf(
  bf(val | resp_se(se, sigma = TRUE) ~ 
       year * sex + year * age +
       (year * sex + year * age |c| country) +
       (year * sex + year * age |r| region),
     sigma ~ 1
  ),
  family = brmsfamily("gaussian", link = "log", link_sigma = "log")
)
```

```{r}
fits_1 <- dat_model1 %>%
  group_by(cause) %>% 
  summarise(
    fit = list(
      brm(
        formula = bf_itu + bf_1 + set_rescor(FALSE), 
        data = cur_data_all(),
        chains = hmc$chains, 
        cores = hmc$cores, 
        threads = hmc$threads,
        iter = hmc$iter, 
        warmup = hmc$warmup, 
        refresh = hmc$refresh,
        control = list(
          adapt_delta = hmc$adapt_delta, 
          max_treedepth = hmc$max_treedepth
        ),
        file = str_glue("output/brm-1-{cur_group()}")
      )
    )
  )

# fits_1 <- dat_model1 %>%
#   group_by(cause) %>% 
#   summarise(
#     fit_se = list(
#       brm(
#         formula = bf_itu + bf_1_se + set_rescor(FALSE), 
#         data = cur_data_all(),
#         chains = hmc$chains, 
#         cores = hmc$cores, 
#         threads = hmc$threads,
#         iter = hmc$iter, 
#         warmup = hmc$warmup, 
#         refresh = hmc$refresh,
#         control = list(
#           adapt_delta = hmc$adapt_delta, 
#           max_treedepth = hmc$max_treedepth
#         ),
#         file = str_glue("output/brm-1-se-{cur_group()}")
#       )
#     )
#   )
```

@@

```{r}
# Create lagged predictors
dat <- dat %>% 
  arrange(region, country, cause, sex, age, year) %>% 
  group_by(region, country, cause, sex, age) %>% 
  mutate(
    across(
      c(internet, fixed, mobile), 
      list(`1` = ~lag(., 1)), 
      .names = "{str_sub(.col, 1, 1)}{.fn}"
    )
  )

# Centering within and between countries
dat <- dat %>%
  group_by(country) %>% 
  mutate(
    i1_cb = mean(i1, na.rm = TRUE), # Country average
    i1_cw = i1 - i1_cb, # Year-specific deviation from country average
    f1_cb = mean(f1, na.rm = TRUE),
    f1_cw = f1 - f1_cb,
    m1_cb = mean(m1, na.rm = TRUE),
    m1_cw = m1 - m1_cb,
  ) %>% 
  ungroup() %>% 
  # Grand-mean center the country averages
  mutate(
    across(c(i1_cb, f1_cb, m1_cb), ~. - mean(., na.rm = T))
  )
```

@@

```{r}
bf_2 <- bf(
  val ~ 
    year * sex + year * age + i1_cw * sex + i1_cw * age + i1_cb + 
    (year * sex + year * age + i1_cw * sex + i1_cw * age | country) +
    (year * sex + year * age + i1_cw * sex + i1_cw * age | region),
  family = gaussian()
)

bf_2_se <- bf(
  bf(val | resp_se(se, sigma = TRUE) ~ 
       year * sex + year * age + i1_cw * sex + i1_cw * age + i1_cb + 
       (year * sex + year * age + i1_cw * sex + i1_cw * age | country) +
       (year * sex + year * age + i1_cw * sex + i1_cw * age | region),
     sigma ~ 1
  ),
  family = brmsfamily("gaussian", link = "log", link_sigma = "log")
)

>>>>

```{r}
itu <- read_rds("data/itu.rds")
gwp <- read_rds("data/gwp.rds")
gbd <- read_rds("data/gbd.rds") 
```

@@

```{r}
# Transform edge proportions to fit beta distribution
# when this is used as an outcome
itu <- itu %>% 
  mutate(ib = if_else(internet == 0, .0000001, internet)) %>% 
  mutate(ib = if_else(internet == 1, .9999999, internet))

# Create lagged predictors
itu <- itu %>% 
  arrange(region, country, year) %>% 
  group_by(region, country) %>% 
  mutate(
    across(
      c(internet, fixed, mobile), 
      list(`1` = ~lag(., 1)), 
      .names = "{str_sub(.col, 1, 1)}{.fn}"
    )
  )

# Centering predictors within and between countries
itu <- itu %>%
  group_by(country) %>% 
  mutate(
    i1_cb = mean(i1, na.rm = TRUE), # Country average
    i1_cw = i1 - i1_cb, # Year-specific deviation from country average
    f1_cb = mean(f1, na.rm = TRUE),
    f1_cw = f1 - f1_cb,
    m1_cb = mean(m1, na.rm = TRUE),
    m1_cw = m1 - m1_cb,
  ) %>% 
  ungroup() %>% 
  # Grand-mean center the country averages
  mutate(
    across(c(i1_cb, f1_cb, m1_cb), ~. - mean(., na.rm = T))
  )
```

```{r data-prepare}
# Stack outcome data
dat <- bind_rows(gwp, gbd)

# Merge predictors to outcomes
dat <- left_join(dat, itu)

# Arrange data
dat <- arrange(dat, region, country, year)

# Center and scale year, and contrast code sex
dat <- dat %>% 
  mutate(year = (year - 2010) / 10)
```

Identify values to use in mv models

```{r}
# Give each outcome value a tag to ensure it is used in MV model
dat <- dat %>% 
  mutate(y = !is.na(val))

# Indicate unique ITU values for MV model
dat <- dat %>% 
  # Unique values exist for each country-year
  group_by(region, country, year) %>% 
  mutate(itu = 1:n() == 1) %>% 
  ungroup()
```

@@

Ensure that factors are contrast coded

```{r}
dat <- dat %>% 
  mutate(across(c(sex, age), factor))
options(contrasts = c(unordered = "contr.sum", ordered = "contr.poly"))
options()$contrasts
```

@@

@@

```{r}
# Define models of outcomes
bf_1 <- bf(
  val ~ 
    year * sex + year * age + 
    (year * sex + year * age |c| country) +
    (year * sex + year * age |r| region),
  family = gaussian(),
  unused = ~ cause
)

bf_1_se <- bf(
  bf(val | resp_se(se, sigma = TRUE) ~ 
       year * sex + year * age +
       (year * sex + year * age |c| country) +
       (year * sex + year * age |r| region),
     sigma ~ 1
  ),
  family = gaussian(),
  unused = ~ cause
)
```

```{r}
fits_1 <- dat %>%
  group_by(cause) %>% 
  summarise(
    fit = list(
      brm(
        formula = bf_itu + bf_1 + set_rescor(FALSE), 
        data = cur_data_all(),
        chains = hmc$chains, 
        cores = hmc$cores, 
        threads = hmc$threads,
        iter = hmc$iter, 
        warmup = hmc$warmup, 
        refresh = hmc$refresh,
        control = list(
          adapt_delta = hmc$adapt_delta, 
          max_treedepth = hmc$max_treedepth
        ),
        file = str_glue("output/brm-1-{cur_group()}")
      )
    )
  )

# fits_1_se <- dat %>%
#   filter(!(cause %in% c("Anxiety", "Depression", "Selfharm")))
#   group_by(cause) %>%
#   summarise(
#     fit_se = list(
#       brm(
#         formula = bf_itu + bf_1_se + set_rescor(FALSE),
#         data = cur_data_all(),
#         chains = hmc$chains,
#         cores = hmc$cores,
#         threads = hmc$threads,
#         iter = hmc$iter,
#         warmup = hmc$warmup,
#         refresh = hmc$refresh,
#         control = list(
#           adapt_delta = hmc$adapt_delta,
#           max_treedepth = hmc$max_treedepth
#         ),
#         file = str_glue("output/brm-1-se-{cur_group()}")
#       )
#     )
#   )
```

@@

@@

```{r}
bf_2 <- bf(
  val ~ 
    year * sex + year * age + i1_cw * sex + i1_cw * age + i1_cb + 
    (year * sex + year * age + i1_cw * sex + i1_cw * age | country) +
    (year * sex + year * age + i1_cw * sex + i1_cw * age | region),
  family = gaussian(),
  unused = ~ cause
)

bf_2_se <- bf(
  bf(val | resp_se(se, sigma = TRUE) ~ 
       year * sex + year * age + i1_cw * sex + i1_cw * age + i1_cb + 
       (year * sex + year * age + i1_cw * sex + i1_cw * age | country) +
       (year * sex + year * age + i1_cw * sex + i1_cw * age | region),
     sigma ~ 1
  ),
  family = gaussian(),
  unused = ~ cause
)
